---
title: "Problem set 3: Global CO~2~"
author: "Your name here"
date: "today's date here"
output: html_document
---

<!--- head ------->
<head>
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
<!--- link rel="icon" 
      type="image/png" 
      href = "www/hex_icon.png" / -->
<!-- link rel="stylesheet" href="https://maxcdn.bootstrapcdn.com/bootstrap/3.3.7/css/bootstrap.min.css" -->
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
<script src = 'https://kit.fontawesome.com/03064bdd6c.js' crossorigin='anonymous'></script>
<!-- link rel="stylesheet" href="https://maxcdn.bootstrapcdn.com/bootstrap/3.3.7/css/bootstrap.min.css" -->
</head>

<hr>

```{css, echo = FALSE}
/* I can add custom classes with a period symbol. See usage of this class below */
.score {
  font-weight: bold;
  font-size: 16px;
  color: red;
  background-color: #ffff80;
}

.subscore {
  font-size: 14px;
  color: red;
}

/* Accordion style, the accordion is the section "Functions that you may use in this assignment" */
  
.accordion {
  background-color: #e6f0ff;
  color: #000000;
  cursor: pointer;
  padding: 18px;
  width: 100%;
  border-style: solid;
  border-width: 2px;
  border-color: #ffffff;
  text-align: left;
  outline: none;
  font-size: 15px;
  transition: 0.4s;
}

/* When you hover your mouse over the accordion button, it changes color. */

.active, .accordion:hover {
  background-color: #80b3ff; 
  color: #ffffff;
  font-weight: bold;
}

/* This is where the content of the accordion is placed. */

.panel {
  padding: 0 18px;
  display: none;
  background-color: #eeffff;
  overflow: hidden;
}
```

## Overview

In this exercise, you will explore data compiled by the <a href = "https://databank.worldbank.org/home.aspx" target = "_blank">World Bank DataBank</a> that provides information on the emissions of greenhouse gases and the populations of countries around the world.

Please start by knitting this file. You can use the html file that is generated as a handy reference (it's much easier to read than the markdown file).

<span style = "color: red; font-size: 18px"><i class="fas fa-exclamation-triangle"></i> **Important!**</span> For R Markdown to work properly this file must be placed in the top level of your data_sci folder.

You will be working in two files as you complete this problem_set, your own new script file, and this R Markdown file. I strongly suggest working in your R script file and copying-and-pasting your code into this R Markdown file, where prompted. 

<hr>

## Grading

The points allotted for each question are provided in highlighted red bold text (e.g., <span class = "score">[1.0]</span>) within the question itself. When applicable, total points for a question may represent the sum of individually graded components, which are provided in red text (e.g., <span class = "subscore">[1.0]</span>).

Points may be deducted from each question's total: 

* Use only functions listed in "Functions that you may use in this assignment" [50%]
* Include only assignments specified in the question [50%]
* Unless specified otherwise, each question must be answered in a single piped statement [50%]
* Include a comment at the start of each code block [25%]
* Include a comment for each line of code that modifies a dataset [10% per violation]
* Ensure that all code is properly indented [10% per violation]
* Ensure that your code follows modern coding conventions (see *Best practices in modern R coding*) [10% per violation]

*Note: The maximum deduction is the total points value for a given question*

You must ensure that your R Markdown document **runs out-of-the-box** -- in other words, the document will knit without error. Some tips for doing so:

* Do not maintain your project folder in a location that is backed up by an online directory (e.g., Dropbox, iCloud)
* Ensure that all file paths are equivalent to those in this document (e.g., in `source()` or `read_csv()`)
* Do not use `setwd()` in your code (***Never** use `setwd()`!*)
* If one of your code chunks contains an error, include `eval = FALSE` in the options for that chunk
* Prior to submission, clear all objects from your global environment prior to knitting your R Markdown document
<hr>

<!-- Note, the below is an accordion, which was styled in the css section -->

<div>
<button class="accordion">Functions that you may use in this assignment</button>
<div class="panel">

In this assignment, you may use **only** the following R functions (*Note: If you are unclear on what a given function does, use `?` to view the help file!*):

* %>%
* %in%
* aes
* arrange
* as.numeric
* bind_cols
* c
* desc
* element_line
* element_rect
* facet_wrap
* filter
* geom_line
* geom_point
* ggplot
* kable
* labs
* inner_join
* left_join
* library
* mutate
* pivot_longer
* pull
* read_csv
* rename
* scale_x_continuous
* scale_y_continuous
* select
* semi_join
* slice
* slice_head
* slice_max
* slice_min
* source
* theme

</div>
</div>

<hr>

## Getting started

Open R Studio. Remember that it's best practice to start with a clean R Studio session (i.e., close any script files, remove objects from your global environment, and clear your history).

1\. <span class = "score">[0.5 points]</span> Save and knit this document:

* <span class = "subscore">[0.1 points]</span> Replace my name in the YAML header with yours
* <span class = "subscore">[0.1 points]</span> Add the current date in the YAML header
* <span class = "subscore">[0.3 points]</span> Save the *.rmd* in the **top level of your project folder** as (*but replace my name with yours*): `problem_set_week_2_Evans_Brian.rmd` 

<hr>

## Read in and tidy the data

I've provided three data sets:

* Country-level information: `world_bank_countries.csv`
* Population data, by country, from 1960 through 2020: `API_SP.POP.TOTL_DS2_en_csv_v2_2763937.csv`
* CO~2~ emission data, by country, from 1960 through 2020: `API_EN.ATM.CO2E.PC_DS2_en_csv_v2_2764620.csv`

*Note: The data set `world_bank_countries.csv` has been cleaned for you. The population and CO~2~ datasets are in their raw form (including the names).*

2\. <span class = "score">[0.5 points]</span> Load the tidyverse library:

```{r q02, message = FALSE, warning = FALSE}
# Paste your code in here!
```


3\. <span class = "score">[1 point]</span>  The following code reads in `world_bank_countries.csv` and assigns the name `countries` to your global environment. Please modify this code such that it follows best management practices in code formatting.

```{r q03, message = FALSE, warning = FALSE}
countries=read_csv("data/raw/world_bank_countries.csv")%>%select(!country_region)
```

4\. <span class = "score">[1 point]</span> The code below reads in and wrangles the data successfully, but not parsimoniously. Modify the code chunk such that the process below is completed in a chained analysis and only the final object, `world_pop`, is assigned to your global environment.

```{r q04, message = FALSE, warning = FALSE}
# Read in population data:

world_pop_raw <-
  read_csv(
    'data/raw/API_SP.POP.TOTL_DS2_en_csv_v2_2763937.csv', 
    skip = 3)

# Subset to columns of interest:

world_pop_column_subset <-
  select(
    world_pop_raw,
    
    # Rename Country Code column to remove space:
    
    country_code = `Country Code`,
    
    # Subset to other columns of interest:
    
    `1960`:`2020`)


# Make the data tidy by pivoting to a long form table:

world_pop_long <-
  pivot_longer(
    world_pop_column_subset,
    names_to = 'year',
    values_to = 'population',
    `1960`:`2020`)

# Finally, convert the year column to numeric:

world_pop <-
  mutate(
    world_pop_long,
    year = as.numeric(year))
```

*Note: Why did I include "skip = 3" as one of my arguments? You might want to look at what happens when you don't include a skip argument. If this is confusing, please revisit the video lecture and tutorial on reading and writing data.*

*Also note: Back ticks (e.g.,* `1960`*) are needed when columns start with a number or contain a space.*

5\. <span class = "score">[1 point]</span> In a single chained analysis:

* Read in the CO~2~ data (`API_EN.ATM.CO2E.PC_DS2_en_csv_v2_2764620.csv`) <span class = "subscore">[0.2 points]</span>
* Wrangle the data such that the resultant tibble contains (only) the variables `country_code`, `year`, and `co2` (as above) <span class = "subscore">[0.2 points]</span>
* Convert the class of the year variable to numeric <span class = "subscore">[0.2 points]</span>
* Assign the data frame to your global environment with the name `co2` <span class = "subscore">[0.2 points]</span>
* Include comments describing each step <span class = "subscore">[0.2 points]</span>

```{r q05, message = FALSE, warning = FALSE}
# Paste your code in here!
```

We can see that the data are not tidy. Recall the tidy rule that every level of observation forms a table. The population table and the CO~2~ table are the same level of observation -- each characterize some attribute of the country on a given year. As such, the data should be stored in the same table.  We'll need to join the tables together.

6\. <span class = "score">[1.5 points]</span> Join the population and CO~2~ tibbles. Assign the name, `population_co2` to the resultant object. In the code chunk below, paste the code section that you used to complete this process. *Hint: You’ll have to be careful here, because the two tibbles contain two common columns that define the observation! This is what is known as a **compound key**. See `?join` for how to address this.*

```{r q06, message = FALSE, warning = FALSE}
# Paste your code in here!
```

<hr>

## Data output

Our final goal will be to visualize the change in per capita CO~2~ emissions from 1960-2018 for the countries that emitted the most total CO~2~ in 2018 (the last year for which we have emissions data). *Note: CO~2~ data are provided in per capita emissions -- metric tons of CO~2~ per person.* 

7\. <span class = "score">[2 points]</span> The code below calculates the total CO~2~ emissions for each country in the `populations_co2` data frame. Modify this code such that the resultant table:

* <span class = "subscore">[0.4 points]</span> Is arranged from the highest to lowest CO~2~ total emissions in 2018 
* <span class = "subscore">[0.4 points]</span> Contains **only** countries with records in the `countries` data frame
* <span class = "subscore">[0.4 points]</span> Contains **only** the 5 countries with the most CO~2~ emissions (and also satisfy the condition above)
* <span class = "subscore">[0.4 points]</span> Contains **only** the fields `country_code`, `country_name`, and `total_co2`
* <span class = "subscore">[0.3 points]</span> Is returned as a `kable` (*Note: `kable()` is a function in the `knitr` package*)
* <span class = "subscore">[0.1 points]</span> In the chunk options, change `eval = FALSE` to `eval = TRUE`

```{r q07, message = FALSE, warning = FALSE, eval = FALSE}
population_co2 %>%
  
  # Subset to 2018:
  
  filter(year == 2018) %>% 
  
  # Calculate the total co2 emissions, per country, in 2018:
  
  mutate(total_co2 = population*co2)
```

8\. <span class = "score">[3 points]</span>  Using the code you generated above (except for the `kable`) generate a ggplot that shows the **per capita** CO~2~ emissions from 1960 to 2018 for the five countries with the highest **total** CO~2~ emissions (*Hint: If you were able to solve Question 9, your data are **almost** ready to be plotted, but not quite!*). Plot the data such that:

* <span class = "subscore">[0.3 points]</span> The data being plotted are the per capita emissions from 1960-2018 for the top five emitting countries
* <span class = "subscore">[0.3 points]</span> Your x-aesthetic is year and is labeled "Year"
* <span class = "subscore">[0.3 points]</span> Your y-aesthetic is co_2 and is labeled "Carbon dioxide emissions (metric tons/person)"
* <span class = "subscore">[0.3 points]</span> Your color aesthetic is country_name and is labeled "Country" (*Hint: It's often easier to modify the data going into a `ggplot` than inside of a `ggplot` function*)
* <span class = "subscore">[0.3 points]</span> Your plot includes point and line geometries
* <span class = "subscore">[0.3 points]</span> The x-axis ranges from 1960 to 2020
* <span class = "subscore">[0.3 points]</span> The y-axis ranges from 0 to 30
* <span class = "subscore">[0.3 points]</span> The plot includes a descriptive title
* <span class = "subscore">[0.3 points]</span> The background of the plot is white and contains no grid lines
* <span class = "subscore">[0.3 points]</span> The axis lines are black


```{r q08, message = FALSE, warning = FALSE}
# Paste your code in here!
```

<hr>

## Submission

Please save your R markdown (.Rmd) and submit (only) your R markdown (.Rmd) file to Blackboard. *Do not submit the html file!*. After submitting, save the R Markdown file and knitted html to your `output` folder.

<hr>

<!-- Note: The below is the javascript that I use to make the accordion button work -->

<script>
var acc = document.getElementsByClassName("accordion");
var i;

for (i = 0; i < acc.length; i++) {
  acc[i].addEventListener("click", function() {
    this.classList.toggle("active");
    var panel = this.nextElementSibling;
    if (panel.style.display === "block") {
      panel.style.display = "none";
    } else {
      panel.style.display = "block";
    }
  });
}
</script>

