---
title: "Problem set grade summaries"
format:
  html:
    embed-resouces: true
    self-contained: true
    standalone: true
    toc: false
    number-sections: false
    theme: 
      - cosmo
      - ../../custom_style.scss
editor: source
editor_options: 
  chunk_output_type: console
---

<button class="accordion">Problem set 1</button>

::: panel

```{r, warning = FALSE, message = FALSE, echo = FALSE}
source("problem_set_grader.R")

gradebook <-
  file.path(
    "https://docs.google.com/spreadsheets/d",
    "15NY5WQjI9Xr_j9TalGZsGjj6sTF8inpbEkNMCcXLizc"
  ) %>% 
  googlesheets4::read_sheet()

```

## Problem set 1

:::{style="background-color: white; font-size: 14px;"}

```{r, message = FALSE, echo = FALSE, warning = FALSE}
grade_summaries_1 <-
  get_grade_summaries(problem_set = 1)
```

Descriptive statistics of scores for Problem set 1:

* Mean: `r grade_summaries_1$mean`
* Median: `r grade_summaries_1$median`
* Standard deviation: `r grade_summaries_1$sd`
* Standard error: `r grade_summaries_1$se`

Here is the distribution of scores:

```{r, warning = FALSE, message = FALSE, echo = FALSE, fig.asp = 0.4}
grade_summaries_1 %>% 
  pluck("grades") %>% 
  ggplot() +
  aes(x = grade) +
  geom_histogram(
    fill = "#dcdcdc",
    color = "black"
  ) +
  scale_x_continuous(
    limits = c(0, 10.1),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 6),
    expand = c(0, 0)
  ) +
  theme_classic()
```

### Points off by question

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_1 %>% 
  pluck("points_off_by_question") %>% 
  mutate(
    ave_percent_off = round(ave_percent_off, 2)
  ) %>% 
  rename(
    Question = question,
    `Average point deduction (%)` = ave_percent_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by the type of mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_1 %>% 
  pluck("points_off_by_mistake_class") %>% 
   mutate(
    ave_points_off = round(ave_points_off, 2)
  ) %>% 
  rename(
    `Type of mistake` = mistake_class,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points of by mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_1 %>% 
  pluck("points_off_by_mistake_description") %>% 
  rename(
    `Mistake` = mistake_description,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

:::
:::

<button class="accordion">Problem set 2</button>

::: panel

## Problem set 2

:::{style="background-color: white; font-size: 14px;"}

```{r, message = FALSE, echo = FALSE, warning = FALSE}
grade_summaries_2 <-
  get_grade_summaries(problem_set = 2)
```

Descriptive statistics of scores for Problem set 2:

* Mean: `r grade_summaries_2$mean`
* Median: `r grade_summaries_2$median`
* Standard deviation: `r grade_summaries_2$sd`
* Standard error: `r grade_summaries_2$se`

Here is the distribution of scores:

```{r, warning = FALSE, message = FALSE, echo = FALSE, fig.asp = 0.4}
grade_summaries_2 %>% 
  pluck("grades") %>% 
  ggplot() +
  aes(x = grade) +
  geom_histogram(
    fill = "#dcdcdc",
    color = "black"
  ) +
  scale_x_continuous(
    limits = c(0, 10.1),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 6),
    expand = c(0, 0)
  ) +
  theme_classic()
```

### Points off by question

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_2 %>% 
  pluck("points_off_by_question") %>% 
  mutate(
    ave_percent_off = round(ave_percent_off, 2)
  ) %>% 
  rename(
    Question = question,
    `Average point deduction (%)` = ave_percent_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by the type of mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_2 %>% 
  pluck("points_off_by_mistake_class") %>% 
   mutate(
    ave_points_off = round(ave_points_off, 2)
  ) %>% 
  rename(
    `Type of mistake` = mistake_class,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points of by mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_2 %>% 
  pluck("points_off_by_mistake_description") %>% 
  rename(
    `Mistake` = mistake_description,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

:::
:::

<button class="accordion">Problem set 3</button>

::: panel

## Problem set 3

:::{style="background-color: white; font-size: 14px;"}

```{r, message = FALSE, echo = FALSE, warning = FALSE}
grade_summaries_3 <-
  get_grade_summaries(problem_set = 3)
```

Descriptive statistics of scores for Problem set 3:

* Mean: `r grade_summaries_3$mean`
* Median: `r grade_summaries_3$median`
* Standard deviation: `r grade_summaries_3$sd`
* Standard error: `r grade_summaries_3$se`

Here is the distribution of scores:

```{r, warning = FALSE, message = FALSE, echo = FALSE, fig.asp = 0.4}
grade_summaries_3 %>% 
  pluck("grades") %>% 
  ggplot() +
  aes(x = grade) +
  geom_histogram(
    fill = "#dcdcdc",
    color = "black"
  ) +
  scale_x_continuous(
    limits = c(0, 10.1),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 6),
    expand = c(0, 0)
  ) +
  theme_classic()
```

### Points off by question

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_3 %>% 
  pluck("points_off_by_question") %>% 
  mutate(
    ave_percent_off = round(ave_percent_off, 2)
  ) %>% 
  rename(
    Question = question,
    `Average point deduction (%)` = ave_percent_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by the type of mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_3 %>% 
  pluck("points_off_by_mistake_class") %>% 
   mutate(
    ave_points_off = round(ave_points_off, 2)
  ) %>% 
  rename(
    `Type of mistake` = mistake_class,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_3 %>% 
  pluck("points_off_by_mistake_description") %>% 
  rename(
    `Mistake` = mistake_description,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```
<button class="accordion">Gradebook</button>

::: panel
:::{style="background-color: white; font-size: 14px;"}

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_3 %>% 
  pluck("grades") %>% 
  arrange(student_name_last) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

:::
:::


:::

:::

<button class="accordion">Problem set 4</button>

::: panel

## Problem set 4

:::{style="background-color: white; font-size: 14px;"}

```{r, message = FALSE, echo = FALSE, warning = FALSE}
grade_summaries_4 <-
  get_grade_summaries(problem_set = 4)
```

Descriptive statistics of scores for Problem set 4:

* Mean: `r grade_summaries_4$mean`
* Median: `r grade_summaries_4$median`
* Standard deviation: `r grade_summaries_4$sd`
* Standard error: `r grade_summaries_4$se`

Here is the distribution of scores:

```{r, warning = FALSE, message = FALSE, echo = FALSE, fig.asp = 0.4}
grade_summaries_4 %>% 
  pluck("grades") %>% 
  ggplot() +
  aes(x = grade) +
  geom_histogram(
    fill = "#dcdcdc",
    color = "black"
  ) +
  scale_x_continuous(
    limits = c(0, 10.1),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 6),
    expand = c(0, 0)
  ) +
  theme_classic()
```

### Points off by question

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_4 %>% 
  pluck("points_off_by_question") %>% 
  mutate(
    ave_percent_off = round(ave_percent_off, 2)
  ) %>% 
  rename(
    Question = question,
    `Average point deduction (%)` = ave_percent_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by the type of mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_4 %>% 
  pluck("points_off_by_mistake_class") %>% 
   mutate(
    ave_points_off = round(ave_points_off, 2)
  ) %>% 
  rename(
    `Type of mistake` = mistake_class,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_4 %>% 
  pluck("points_off_by_mistake_description") %>% 
  rename(
    `Mistake` = mistake_description,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```
<button class="accordion">Gradebook</button>

::: panel
:::{style="background-color: white; font-size: 14px;"}

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_4 %>% 
  pluck("grades") %>% 
  arrange(student_name_last) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

:::
:::

:::
:::

<button class="accordion">Problem set 5</button>

::: panel

## Problem set 5

:::{style="background-color: white; font-size: 14px;"}

```{r, message = FALSE, echo = FALSE, warning = FALSE}
grade_summaries_5 <-
  get_grade_summaries(problem_set = 5)
```

Descriptive statistics of scores for Problem set 5:
  
* Mean: `r grade_summaries_5$mean`
* Median: `r grade_summaries_5$median`
* Standard deviation: `r grade_summaries_5$sd`
* Standard error: `r grade_summaries_5$se`

Here is the distribution of scores:
  
```{r, warning = FALSE, message = FALSE, echo = FALSE, fig.asp = 0.4}
grade_summaries_5 %>% 
  pluck("grades") %>% 
  ggplot() +
  aes(x = grade) +
  geom_histogram(
    fill = "#dcdcdc",
    color = "black"
  ) +
  scale_x_continuous(
    limits = c(0, 10.3),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 6),
    expand = c(0, 0)
  ) +
  theme_classic()
```

### Points off by question

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_5 %>% 
  pluck("points_off_by_question") %>% 
  mutate(
    ave_percent_off = round(ave_percent_off, 2)
  ) %>% 
  rename(
    Question = question,
    `Average point deduction (%)` = ave_percent_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by the type of mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_5 %>% 
  pluck("points_off_by_mistake_class") %>% 
  mutate(
    ave_points_off = round(ave_points_off, 2)
  ) %>% 
  rename(
    `Type of mistake` = mistake_class,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_5 %>% 
  pluck("points_off_by_mistake_description") %>% 
  rename(
    `Mistake` = mistake_description,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

<button class="accordion">Gradebook</button>
  
::: panel
:::{style="background-color: white; font-size: 15px;"}

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_5 %>% 
  pluck("grades") %>% 
  arrange(student_name_last) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

:::
:::
  
:::
:::

<button class="accordion">Problem set 6</button>

::: panel

## Problem set 6

:::{style="background-color: white; font-size: 14px;"}

```{r, message = FALSE, echo = FALSE, warning = FALSE}
grade_summaries_6 <-
  get_grade_summaries(problem_set = 6)
```

Descriptive statistics of scores for Problem set 6:
  
* Mean: `r grade_summaries_6$mean`
* Median: `r grade_summaries_6$median`
* Standard deviation: `r grade_summaries_6$sd`
* Standard error: `r grade_summaries_6$se`

Here is the distribution of scores:
  
```{r, warning = FALSE, message = FALSE, echo = FALSE, fig.asp = 0.4}
grade_summaries_6 %>% 
  pluck("grades") %>% 
  ggplot() +
  aes(x = grade) +
  geom_histogram(
    fill = "#dcdcdc",
    color = "black"
  ) +
  scale_x_continuous(
    limits = c(0, 10.3),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 6),
    expand = c(0, 0)
  ) +
  theme_classic()
```

### Points off by question

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_6 %>% 
  pluck("points_off_by_question") %>% 
  mutate(
    ave_percent_off = round(ave_percent_off, 2)
  ) %>% 
  rename(
    Question = question,
    `Average point deduction (%)` = ave_percent_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by the type of mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_6 %>% 
  pluck("points_off_by_mistake_class") %>% 
  mutate(
    ave_points_off = round(ave_points_off, 2)
  ) %>% 
  rename(
    `Type of mistake` = mistake_class,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_6 %>% 
  pluck("points_off_by_mistake_description") %>% 
  rename(
    `Mistake` = mistake_description,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

<button class="accordion">Gradebook</button>
  
::: panel
:::{style="background-color: white; font-size: 16px;"}

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_6 %>% 
  pluck("grades") %>% 
  arrange(student_name_last) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

:::
:::
  
:::
:::

<button class="accordion">Problem set 7</button>

::: panel

## Problem set 7

:::{style="background-color: white; font-size: 14px;"}

```{r, message = FALSE, echo = FALSE, warning = FALSE}
grade_summaries_7 <-
  get_grade_summaries(problem_set = 7)
```

Descriptive statistics of scores for Problem set 7:
  
* Mean: `r grade_summaries_7$mean`
* Median: `r grade_summaries_7$median`
* Standard deviation: `r grade_summaries_7$sd`
* Standard error: `r grade_summaries_7$se`

Here is the distribution of scores:
  
```{r, warning = FALSE, message = FALSE, echo = FALSE, fig.asp = 0.4}
grade_summaries_7 %>% 
  pluck("grades") %>% 
  ggplot() +
  aes(x = grade) +
  geom_histogram(
    fill = "#dcdcdc",
    color = "black"
  ) +
  scale_x_continuous(
    limits = c(0, 11),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 7),
    expand = c(0, 0)
  ) +
  theme_classic()
```

### Points off by question

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_7 %>% 
  pluck("points_off_by_question") %>% 
  mutate(
    ave_percent_off = round(ave_percent_off, 2)
  ) %>% 
  rename(
    Question = question,
    `Average point deduction (%)` = ave_percent_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by the type of mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_7 %>% 
  pluck("points_off_by_mistake_class") %>% 
  mutate(
    ave_points_off = round(ave_points_off, 2)
  ) %>% 
  rename(
    `Type of mistake` = mistake_class,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

### Points off by mistake

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_7 %>% 
  pluck("points_off_by_mistake_description") %>% 
  rename(
    `Mistake` = mistake_description,
    `Average points off` = ave_points_off) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

<button class="accordion">Gradebook</button>
  
::: panel
:::{style="background-color: white; font-size: 17px;"}

```{r, warning = FALSE, message = FALSE, echo = FALSE}
grade_summaries_7 %>% 
  pluck("grades") %>% 
  arrange(student_name_last) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_paper("hover", full_width = F)
```

:::
:::
  
:::
:::


<script>
var acc = document.getElementsByClassName("accordion");
var i;

for (i = 0; i < acc.length; i++) {
acc[i].addEventListener("click", function() {
this.classList.toggle("active");
var panel = this.nextElementSibling;
if (panel.style.display === "block") {
panel.style.display = "none";
} else {
panel.style.display = "block";
}
});
}
</script>

